---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(a11ytables)
```

# {a11ytables}

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![CRAN
status](https://www.r-pkg.org/badges/version/a11ytables)](https://CRAN.R-project.org/package=a11ytables)
[![R-CMD-check](https://github.com/co-analysis/a11ytables/workflows/R-CMD-check/badge.svg)](https://github.com/co-analysis/a11ytables/actions)
[![Codecov test
coverage](https://codecov.io/gh/co-analysis/a11ytables/branch/main/graph/badge.svg)](https://codecov.io/gh/co-analysis/a11ytables?branch=main)
<!-- badges: end -->

## Purpose

{a11ytables} is a work-in-progress R package built around [{openxlsx}](https://ycphs.github.io/openxlsx/) that aims to comply with the latest guidance (June 2021) on [releasing statistics in spreadsheets](https://gss.civilservice.gov.uk/policy-store/releasing-statistics-in-spreadsheets/) from the [Best Practice and Impact Division](https://github.com/best-practice-and-impact?language=html) (BPID) of the UK's [Government Statistical Service](https://gss.civilservice.gov.uk/).

## Install and use

You can install {a11ytables} from GitHub using {remotes}. {openxlsx} will be installed too.

```{r install, eval=FALSE}
install.packages("remotes")
remotes::install_github("co-analysis/a11ytables")
library(a11ytables)
```

To use:

1. Create an a11ytables-class dataframe object containing your data (TODO: make the class)
2. Pass that object to `create_a11y_wb()` to populate an {openxlsx} Workbook-class object
3. Write that object with `openxlsx::saveWorkbook()`

## Detail

The information required to build the output can be arranged into a single dataframe with class 'a11ytables' (TODO: explain how to construct these objects). You pass that object to the `create_a11y_wb()` function, which in turn creates an {openxlsx} Workbook-class object with the required sheets, their content and styling in accordance with accessibility best practice (TODO: make all features accessible).

Here is the built-in dataset `lfs_tables`, adapted from [the releasing statistics in spreadsheets guidance](https://gss.civilservice.gov.uk/policy-store/releasing-statistics-in-spreadsheets/) (TODO: actually convert this a11ytables-class):

```{r glimpse-lfs-tables}
dplyr::glimpse(tibble::as_tibble(lfs_tables))
```

<details><summary>Click here to view an explanation of the variables.</summary>

Columns in the a11ytables-class dataframe are:

* `tab_title`: the name of the tab (i.e. the text that will appear on each tab of the output workbook), a column that must include 'cover' and 'contents'; optionally 'notes'; and one row per table for publication (numbers are recommended for tables)
* `sheet_type`: 'cover', 'contents' and 'notes' are 'meta' sheets; tables for publication are 'tables'
* `sheet_title`: the title of the sheet, displayed in cell A1
* `source`: the provenance of the data (optional), displayed in cell A3
* `table_name`: a name to give the marked-up table (no spaces), which will allow screenreaders to identify the table in the output workbook
* `table`: a list-column containing a dataframe with the table to be added to  a given sheet

Optionally, for tabs that contain more than one table (not recommended, also TODO):

* `subtable_num`: string to append to the tab_title to identify this sub-table, for example 'a' and 'b'
* `subtable_title`: title to be added above the sub-table within the sheet
</details>
<p>

And so, this can be passed to a single function to create an {openxlsx} Workbook-class object:

```{r example_wb}
example_wb <- create_a11y_wb(lfs_tables)
```

<details><summary>Click here to view the contents of the output.</summary>

```{r}
example_wb
```

</details>
<p>

This output can be written to disk or opened temporarily with spreadsheet software.

```{r open-wb, eval=FALSE}
openxlsx::openXL(example_wb)  # optionally open temp copy
openxlsx::saveWorkbook(example_wb, "publication.xlsx")
```

## Similar projects

BPID have themselves released [a Python package called 'gptables'](https://github.com/best-practice-and-impact/gptables) to help users produce such spreadsheets, but it's based on a prior version of the guidance. [That package will be updated](https://github.com/best-practice-and-impact/gptables/issues/145), but there's a gap for this functionality—and for a dedicated R package—at the time of writing (August 2021).

{a11ytables} is being developed with specific statistical releases in mind, but the goal is to generalise it and perhaps have some functionality absorbed into the development of [gptables](https://github.com/best-practice-and-impact/gptables) in some way.

The code is under development and current API features may change in future releases. Please see the NEWS file for details of any breaking changes.

## Code of Conduct

Please note that the {a11ytables} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

Cabinet Office users, please adhere to [the co-analysis Git and GitHub code of conduct and usage policy](https://docs.google.com/document/d/1CuNgKla1BwSVOmGkPmsq0S-OM4emP-iXrgnm7EeILWM/edit?usp=sharing).

Please [file an issue](https://github.com/co-analysis/csstatsbulletin/issues) where possible and provide meaningful commentary in any commits and pull requests.

## Copyright

Copyright (c) 2021 Crown Copyright (Cabinet Office).

The code in this repository is released under the MIT license as per [The GDS Way](https://gds-way.cloudapps.digital/manuals/licensing.html#use-mit).
