---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(a11ytables)
library(pillar)
```

# {a11ytables}

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![CRAN
status](https://www.r-pkg.org/badges/version/a11ytables)](https://CRAN.R-project.org/package=a11ytables)
[![R-CMD-check](https://github.com/co-analysis/a11ytables/workflows/R-CMD-check/badge.svg)](https://github.com/co-analysis/a11ytables/actions)
[![Codecov test
coverage](https://codecov.io/gh/co-analysis/a11ytables/branch/main/graph/badge.svg)](https://codecov.io/gh/co-analysis/a11ytables?branch=main)
<!-- badges: end -->

## Purpose

{a11ytables} is a work-in-progress R package built around [{openxlsx}](https://ycphs.github.io/openxlsx/) that aims to help you comply with the latest guidance (June 2021) on [releasing statistics in spreadsheets](https://gss.civilservice.gov.uk/policy-store/releasing-statistics-in-spreadsheets/) from the [Best Practice and Impact Division](https://github.com/best-practice-and-impact?language=html) (BPID) of the UK's [Government Statistical Service](https://gss.civilservice.gov.uk/).

Visit [the {a11ytables} website](https://co-analysis.github.io/a11ytables/) for documentation.

The package is under development. Please see [the NEWS file](https://co-analysis.github.io/a11ytables/news/index.html) for changes.

## How to use

You can install {a11ytables} from GitHub using {remotes}. The {openxlsx} and {pillar} packages will be installed too.

```{r install, eval=FALSE}
install.packages("remotes")  # if not already installed
remotes::install_github("co-analysis/a11ytables")
library(a11ytables)
```

The package is designed with simplicity and minimal user input in mind. There are only three steps to creating your output workbook:

1. Create an a11ytable-class object by either:
    a. passing data as arguments to `new_a11ytable()` (recommended)
    b. creating a dataframe in the correct format and coercing it with `as_a11ytable()`
1. Pass your a11ytables-class object to `create_a11y_wb()` to generate an {openxlsx} Workbook-class object with the required styling
1. Write the workbook to a spreadsheet output with `openxlsx::saveWorkbook()`

Visit [the 'construct an a11ytable' vignette](https://co-analysis.github.io/a11ytables/articles/construct.html) to see how to set up an a11ytable-class object. Visit [the 'accessibilty guidance checklist' vignette](https://co-analysis.github.io/a11ytables/articles/accessibility.html) to see how well the package adheres [to best practice](https://gss.civilservice.gov.uk/policy-store/making-spreadsheets-accessible-a-brief-checklist-of-the-basics/).

## Example

### 1. Create a11ytable object

First, use `new_a11ytable()` to construct a special a11ytable-class object, which is just a data.frame with strict requirements. It must have a row per sheet, specific columns for meta information (e.g. sheet titles, data sources) and a list-column of data.frames that contain the actual data; run `?new_a11ytable` for details.

There's a built-in demo dataset---`mtcars_df`, adapted from R's built-in mtcars demo dataset---that is already in the correct data.frame format with the appropriate variables and content.

```{r glimpse-mtcars-tables}
str(mtcars_df, max.level = 2)
```

It can be coerced to a11ytable class with `as_a11ytable()`, since it's already structured correctly.

```{r as-a11ytable}
mtcars_a11ytable <- as_a11ytable(mtcars_df)
```

The `new_a11ytable()` and `as_a11ytable()` functions will run validation on the generated object to ensure it meets the requirements of the a11ytable class.

Note that the a11ytable object also has class 'data.frame' and can be manipulated as such.

```{r check-class}
class(mtcars_a11ytable)  # see classes
is_a11ytable(mtcars_a11ytable)  # check if a11ytable class
```

The object also has class 'tbl', so it's pretty-printed in {tibble} style thanks to the imported {pillar} package.

```{r}
mtcars_a11ytable
```

### 2. Convert to workbook

We can pass this a11ytable-class object to `create_a11y_wb()`, which constructs an {openxlsx} Workbook-class object with appropriate styling and markup.

```{r create-a11ytable-wb}
mtcars_wb <- create_a11y_wb(mtcars_a11ytable)
```

<details><summary>Click here to view the contents of the workbook</summary>

```{r print-wb}
mtcars_wb
```

</details>
<p>

### 3. Save

The Workbook-class object can be written to disk or opened temporarily with spreadsheet software.

```{r open-wb, eval=FALSE}
openxlsx::openXL(mtcars_wb)  # optionally open temp copy
openxlsx::saveWorkbook(mtcars_wb, "publication.xlsx")
```

The {a11ytables} packaged is designed for general use and to minimise user interaction. As a result of this interface, the final output may not suit your needs exactly and you may need to tweak it. It's recommended for maximum reproducibility that you record your adjustments as code, rather than changing the content of the output file.

## Similar projects

BPID have themselves released [a Python package called 'gptables'](https://github.com/best-practice-and-impact/gptables) to help users produce such spreadsheets, but it's based on a prior version of the bset practice guidance. [That package will be updated](https://github.com/best-practice-and-impact/gptables/issues/145), but there's a gap for this functionality—and for a dedicated R package—at the time of writing (August 2021).

{a11ytables} is being developed with specific statistical releases in mind, but the goal is to generalise it and perhaps have some functionality absorbed into the development of [gptables](https://github.com/best-practice-and-impact/gptables) in some way.

## Code of Conduct

Please note that the {a11ytables} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

Cabinet Office users, please adhere to [the co-analysis Git and GitHub code of conduct and usage policy](https://docs.google.com/document/d/1CuNgKla1BwSVOmGkPmsq0S-OM4emP-iXrgnm7EeILWM/edit?usp=sharing).

Please [file an issue](https://github.com/co-analysis/csstatsbulletin/issues) where possible and provide meaningful commentary in any commits and pull requests.

## Copyright

Copyright (c) 2022 Crown Copyright (Cabinet Office).

The code in this repository is released under the MIT license as per [The GDS Way](https://gds-way.cloudapps.digital/manuals/licensing.html#use-mit).
